---
title: "12-virmerge"
title-block-banner: true
author:
  - name: Jinlong Ru
    orcid: 0000-0002-6757-6018
toc: true
code-fold: true
format: 
  html:
    embed-resources: true
    smooth-scroll: true
bibliography: refs.bib
params:
  name: "12-virmerge"
---

Updated: `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'CET')` CET.


```{r setup}
#| message: false
#| include: false
here::i_am(paste0(params$name, ".qmd"), uuid = "dde7713c-6130-4ad7-8454-96d7e77ea2fd")
projthis::proj_create_dir_target(params$name, clean = FALSE)
path_target <- projthis::proj_path_target(params$name)
path_source <- projthis::proj_path_source(params$name)
path_data <- path_source("00-raw", params$name)
dir.create(path_data, recursive = T)
```

```{r packages}
#| warning: false
#| message: false
library(conflicted)
library(here)
library(tidyverse)
library(miaViz)

conflict_prefer("setdiff", "dplyr")
```


## Tasks

The purpose of this part is to import TSE with all contigs, and then,

1. Filter viral contigs
2. Create a new TSE with only viral contigs
3. Adjust contig abundance using covfrac information
4. Combine contig annotation with adjusted abundance information and save to excel file.

```{r import}
fin_meta <- path_source("00-raw", "p0094-metadata.xlsx")
fin_tseall <- here(path_data, "viroprofiler_output_all_contigs.rds")

tseall <- readRDS(fin_tseall)
# colData(tseall) <- colData(tseall) %>% 
#   data.frame() %>% 
#   dplyr::left_join(df_meta, by = "sample_name") %>% 
#   column_to_rownames("sample_name") %>% 
#   DataFrame()

# filter Viral contigs
vir_mmseqs <- !is.na(rowData(tseall)$Domain)
vir_checkv <- rowData(tseall)$checkv_quality %in% c("Complete", "High-quality", "Medium-quality")
vir_virsorter2 <- rowData(tseall)$virsorter2_max_score_group %in% c("dsDNAphage", "NCLDV", "RNA", "ssDNA", "lavidaviridae")
vir_dvf <- !is.na(rowData(tseall)$dvf_score)
vir_vibrant <- rowData(tseall)$vibrant_quality %in% c("complete circular", "high quality draft", "medium quality draft")
vir_all <- vir_checkv | vir_virsorter2 | vir_vibrant | vir_dvf

tse <- tseall[vir_all,]
```

Adjust counts by covfrac.

```{r}
create_viral_abundance <- function(df_abundance, df_covfrac, covfrac_threshold=0.5) {
  df_covfrac[df_covfrac<covfrac_threshold] <- 0
  df_covfrac[df_covfrac>=covfrac_threshold] <- 1
  df_abundance_filtered <- df_abundance*df_covfrac
  return(df_abundance_filtered)
}

tbl_counts_adjusted_by_covfrac <- create_viral_abundance(assay(tse, "counts"), assay(tse, "covfrac"), 0.5) %>% 
  rownames_to_column("feature_id")

tbl_report <- rowData(tse) %>% 
  data.frame() %>% 
  rownames_to_column("feature_id") %>% 
  dplyr::left_join(tbl_counts_adjusted_by_covfrac, by = "feature_id")
```


Save to excel file

```{r}
openxlsx::write.xlsx(tbl_report, file = path_target("p0094_virmerge_viroprofiler.xlsx"))
# openxlsx::write.xlsx(tbl_meta, file = path_target("p0094_virmerge_metadata.xlsx"))
```

## Files written

These files have been written to the target directory, ```r paste0("data/", params$name)```:

```{r list-files-target}
projthis::proj_dir_info(path_target())
```
